---
title: "Average Cost of the Most Common Heart Procedure in the U.S by State"
author: "Israel Diego"
date: "`r format.Date(Sys.Date(), '%B %d, %Y')`"
output:
  html_document:
    includes:
      in_header: ./Web_Materials/header.html
      after_body: ./Web_Materials/disqus.html
    toc: true
    toc_float: true
    theme: cerulean
    highlight: pygments
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(ggplot2)
library(tidyr)
library(dplyr)
library(directlabels)
```

## Introduction

We wish to figure out the most common heart procedure in the U.S using the 
*Medicare Physician and Other Supplier Data* for years (2012-2016) and then
calculate the average cost of this procedure for each state over our five-year 
time period. 

## Methods

We extract the following columns for each year (2012-2016) using `dplyr`: 

* **HCPCS_CODE**: Code for the specific medical service
* **HCPCS_DESCRIPTION**: Description of the HCPCS code for the specific medical 
service
  + We will search in the description for all heart procedures
* **NPPES_PROVIDER_STATE**: The state where the provider is located
* **LINE_SRVC_CNT**: Number of Services
  + We will extract the heart procedure with the most number of services
* **AVERAGE_MEDICARE_PAYMENT_AMT**: Average Medicare Payment amount for service


## Results
```{r ps4_q2c_source, message = FALSE, echo = FALSE}
source('./individual_project.R')
```

First we find the top Heart Procedures in the U.S, which are listed below. We
determine that ***Ultrasound Examination of Heart*** is the most common heart 
procedure in the U.S.

```{r results_table1, echo = FALSE}
cap = '**Table 1.** Top 5 Heart Procedures in the U.S'

# Pretty Printing
pwc = function(x) format(round(x), big.mark = ',')

# Display Procedure and Total number of services provided for this procedure
topProcCode %>%
  transmute( 
    `Total Services` = pwc(count),
    `Description` = description) %>%
  knitr::kable( align = 'r', caption = cap)
```

Next, we compute the average cost of this procedure for each U.S state from (2012-2016). For brevity, we will display the top five states with the 
highest average cost over the 5-year period. Likewise, we will display the 
five states with the lowest average cost of this time period. 

```{r results_table2, echo = FALSE}
cap = '**Table 1.** Average annual Heart Ultra Sound Cost'

# Pretty Printing
pwc = function(x) format(round(x, digits = 2), big.mark = ',')

stateRanks %>%
  transmute( 
    `State` = state,
    `2012` = paste0('$', pwc(avg2012)),
    `2013` = paste0('$', pwc(avg2013)),
    `2014` = paste0('$', pwc(avg2014)),
    `2015` = paste0('$', pwc(avg2015)),
    `2016` = paste0('$', pwc(avg2016)),
    `Overall Average` = paste0('$', pwc(avgOverall))) %>%
  knitr::kable( align = 'r', caption = cap)
```

In order to visualize the trends in our time-series we will display the results
graphically.

```{r results_figure, echo = FALSE, fig.cap = cap}
cap = '**Figure 1.** Average annual Cost of Heart Ultra Sound, Top 5 and 
Lowest 5 U.S States'
# Plot the 10 states, top 5 and lowest 5 average cost
ggplot(aggregateStateAvgPayment, 
                       aes(x = year, 
                           y = avgPayment, 
                           group = state, 
                           color = state)) +
  geom_line() +
  geom_dl(aes(label = state), 
          method = list(dl.combine("first.points"), cex = 0.7)) +
  ggtitle("Average Payment for Heart Ultra Sound By State (2012-2016)") +
  ylab('Average Payment for Service ($)' ) +
  theme(plot.title = element_text(hjust = 0.5))
```

## Discussion
First, we find that Ultra Sounds of the Heart are the most commonly used heart
procedure in the U.S. Second, the average cost of an ultra sound decreased
dramatically for all states in 2014. This trend can be further 
[evidenced](https://en.wikipedia.org/wiki/Patient_Protection_and_Affordable_Care_Act) 
by the establishment of the Affordable Care Act (ACA) whose major provisions 
came into force in 2014. 