---
title: "Problem Set 2"
author: "Israel Diego"
date: "October 19, 2018"
output:
  html_document:
    includes:
      in_header: ./Web_Materials/header.html
      after_body: ./Web_Materials/disqus.html
    toc: true
    toc_float: true
    theme: cerulean
    highlight: pygments
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
library(knitr)
library(gdata)
library(dplyr)
library(faraway)
source('ps2_q3.R', local = TRUE, echo = FALSE)
```

## Question 1
Below we show estimates of national totals for residential energy consumption using Stata:

i. Electricity usage in kilowatt hours

i. Natural gas usage, in hundreds of cubic feet

i. Propane usage, in gallons

i. Fuel oil or kerosene usage, in gallons

```{r Table_q1, echo= FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(knitr)

df = gdata::read.xls("ps2_q2_Stata.xlsx", sheet = "q1", header = TRUE)
df = df %>% 
  select(Variables, Total, BRR.Std..Err., X95..Lower, X95..Upper)
df[,2:5] = df[,2:5] / (10^9)
cap_title = '**National Totals for residential energy consumption 2015**'
cap_text0 = 'Results expressed in (Billions)'
cap_text1 = ''
cap = paste(cap_title, cap_text0, cap_text1)
cols = c('', 
         'Totals', 
         'BRR Std Err.',
         '95% Low',
         '95% Upper')
kable(df, caption = cap, col.names = cols, digits = 3)
```

\pagebreak

## Question 2/3
For this question we use the 2005-2006 Nhanes oral health data and 2005-2006 
demographic data. Analysis is done in Stata and in R. 

### Part a. 
First we merge both data sets by participant id. Which drops some unusued 
observations in the demographic data set. Mainly going from 10348 observations 
in the demographic data to 8305 observations in the oral health data. 

### Part b.
Using logistic regression, we estimate the relationship between age (in months) 
and the probability that an individual has lost a primary upper right 2nd 
bicuspid.
Using the fitted model we estimate the ages at which 25, 50, and 75% of 
individuals lose their primary upper right 2nd bicuspid. 

Below are the are the results of our logistic regression.

```{r Table_q2_b, echo = FALSE}
df = gdata::read.xls("ps2_q2_Stata.xlsx", sheet = "partb", header = TRUE)
df = df %>% select(notPrimary, Coef., Std..Err., X95..Lower , X95..Upper)
cap_title = '**Logistic Regression of monthly age on presence of Primary '
cap_text0 = 'Tooth Upper 2nd Bicuspid**'
cap_text1 = ''
cap = paste(cap_title, cap_text0, cap_text1)
cols = c('notPrimary', 
         'Coef.',
         'Std. Err.',
         '95% Low',
         '95% Upper')
kable(df, caption = cap, col.names = cols, digits = 2)
```

On the next page we also plot our fitted model.

```{r Table_q2_b_plot, echo = FALSE}
range = data.frame(ageMonths = seq(0,1000, by = 1))
pred.l = ilogit(predict(fit0, range))
matplot(range, pred.l, xlim = c(0,250),
        ylim = c(0,1), xlab = "Age in Months", 
        ylab = "Prob Primary Tooth is Not Present",
        type = 'l', lty = 'solid', 
        main = "Logistic Regression of monthly \n age on presence of primary tooth",
        col = "blue")
```

Below are the predicted ages in months and in years at which a primary tooth 
would be lost at the 25, 50, 75% levels.

```{r Table_q2_b2, echo = FALSE}
df = gdata::read.xls("ps2_q2_Stata.xlsx", sheet = "Ages", header = TRUE)
cap_title = '**Representative Ages at 25, 50, 75% levels according to our '
cap_text0 = ' fitted model**'
cap_text1 = ''
cap = paste(cap_title, cap_text0, cap_text1)
cols = c('Age (Months) at 25%', 
         'Age (Months) at 50%',
         'Age (Months) at 75%',
         'Age (Years) at 25%',
         'Age (Years) at 75%')
kable(df, caption = cap, col.names = cols, digits = 2)
```

### Part c.
Continuing from the regression in part b, we now control for separate 
demographics, such as: Gender, Ethnicity, and Income Poverty Ratio.

```{r Table_q2_c, echo = FALSE}
df = gdata::read.xls("ps2_q2_Stata.xlsx", sheet = "partc", header = TRUE)
cap_title = '**Various Logistic Regressions on some demographic variables**'
cap_text0 = ''
cap_text1 = ''
cap = paste(cap_title, cap_text0, cap_text1)
cols = c('', 
         'Age',
         'Age/Gender',
         'Age/Mex',
         'Age/Black',
         'Age/Black/Other',
         'Age/Black/InPovRatio')
kable(df, caption = cap, col.names = cols, digits = 2)
```

i. We do not retain gender, because this caused an increase in BIC. 

ii. We do not retain Mexican, or Other categories of ethnicities, but we do 
retain Non-Hispanic Black.

iii. We retain poverty income ratio because it drops BIC to 1462.895. As a final 
result we retain black ethnicity and Income Poverty Ratio to the regression 
which both improved BIC. 

### Part d.
Now we use the margins command to compute the following:

i. Adjusted predctions at the mean (for other values) at each of the 
representative ages (8 - 12 years old).

```{r Table_q2_d1, echo = FALSE}
df = gdata::read.xls("ps2_q2_Stata.xlsx", sheet = "partd1", header = TRUE)
df = df[,c(1:3,6:7)]
df$Rep..Ages = c(8:12)
cap_title = '**Adjusted predictions at the mean**'
cap_text0 = ''
cap_text1 = ''
cap = paste(cap_title, cap_text0, cap_text1)
cols = c('Age (Years)',
         'Margin',
         'Std. Err.',
         '95% Low',
         '95% Upper')
kable(df, caption = cap, col.names = cols, digits = 3)
```

Below we show a plot of the adjusted predictions which demonstrates that the 
representative ages are evenly spaced out.

```{r q2_d1_plot, results = TRUE, echo = FALSE}
nhanes_atmean %>% 
  ggplot(aes(y = ageMonths, x = fit)) +
  geom_point() +
  geom_errorbarh(aes(xmin = lwr, xmax = upr), height = .1) +
  geom_vline(aes(xintercept = fit), lty = 'dashed', color = 'grey', alpha = .5) +
  theme_bw() +
  xlim(c(0,1)) +
  ggtitle("Adjusted Predictions at the mean") +
  theme(plot.title = element_text(hjust = 0.5))
```

\pagebreak

i. The marginal effects at the mean of Black and Income Poverty Ratio variables 
at the representative ages (8 - 12 years old).

```{r Table_q2_d2, echo = FALSE}
df = gdata::read.xls("ps2_q2_Stata.xlsx", sheet = "partd2", header = TRUE)
df = df[,c(1:3,6:7)]
df$Rep..Ages = c(8:12)
cap_title = '**Marginal Effects at the Mean**'
cap_text0 = ''
cap_text1 = ''
cap = paste(cap_title, cap_text0, cap_text1)
cols = c('Age (Years)', 
         'MEM',
         'Std. Err.',
         '95% Low',
         '95% Upper')
kable(df, caption = cap, col.names = cols, digits = 4)
```

i. The average marginal effect of Black and Income Poverty Ratio variables 
at the representative ages (8 - 12 years old).

```{r Table_q2_d3, echo = FALSE}
df = gdata::read.xls("ps2_q2_Stata.xlsx", sheet = "partd3", header = TRUE)
df = df[,c(1:3,6:7)]
df$Rep..Ages = c(8:12)
cap_title = '**Average Marginal Effects at Representative Ages**'
cap_text0 = ''
cap_text1 = ''
cap = paste(cap_title, cap_text0, cap_text1)
cols = c('Age (Years)', 
         'AME',
         'Std. Err.',
         '95% Low',
         '95% Upper')
kable(df, caption = cap, col.names = cols, digits = 4)
```

### Part e.
We refit the final model from part c using svyset and compare the new model to 
the old model.

```{r Table_q2_c2, echo = FALSE}
df = gdata::read.xls("ps2_q2_Stata.xlsx", sheet = "partc2", header = TRUE)
df = df[,c(1:3,5:7)]
cap_title = '**Logit Regression with Before Survey Weighting**'
cap_text0 = ''
cap_text1 = ''
cap = paste(cap_title, cap_text0, cap_text1)
cols = c('notPrimary', 
         'Coef.',
         'Std. Err.',
         'p-value',
         '95% Low',
         '95% Upper')
kable(df, caption = cap, col.names = cols, digits = 4)
```

```{r Table_q2_e, echo = FALSE}
df = gdata::read.xls("ps2_q2_Stata.xlsx", sheet = "parte", header = TRUE)
df = df[,c(1:3,5:7)]
cap_title = '**Logit Regression with Survey Weighting**'
cap_text0 = ''
cap_text1 = ''
cap = paste(cap_title, cap_text0, cap_text1)
cols = c('notPrimary', 
         'Coef.',
         'Std. Err.',
         'p-value',
         '95% Low',
         '95% Upper')
kable(df, caption = cap, col.names = cols, digits = 4)
```

We notice that there are some slight changes in regression coefficients. The 
main result is that income poverty ratio is no longer significant as 
demonstrated by it's p-value of 0.14. However, the age and black ethnicity 
predictors are still highly significant even though their standard errors 
increased a bit. 