---
title: "Homework 1"
author: "Israel Diego"
date: "January 21, 2016"
output:
  html_document:
    includes:
      in_header: ./Web_Materials/header.html
      after_body: ./Web_Materials/disqus.html
    toc: true
    toc_float: true
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

The dataset `teengamb` concerns a study of teenage gambling in Britain.
More details about the study can be found in Ide-Smith & Lea (1988) Journal of 
Gambling Behavior, 4, 110-118. It contains five variables:

- `sex`: 0 = male; 1 = female
- `status`: socioeconomic status score based on parents' occupation
- `income`: in pounds per week
- `verbal`: verbal score in words out of 12 correctly defined
- `gamble`: expenditure on gambling in pounds per year

Make a numerical and graphical summary of the data, commenting on
any features that you find interesting. Limit the output you present to a
quantity that a busy reader would find sufficient to get a basic understanding
of the data. For the variable `sex`, assign labels "male" and "female" and
make sure R is treating it as a categorical variable before you compute the
summary.
Also answer the following questions, **incorporating the answers into your report**:

1. In your summary, report the means and medians of variables "income"
and "gamble". Explain why the means are larger than the medians.
2. How many different values are there for the variable "verbal"? Based
on the boxplot and/or other summaries of the variable "verbal", what
values of the verbal score would you consider to be outliers? Give the
row numbers of the outlier observations.

**Hints:** After you install `R` (if it is not already installed on the system 
you are using) and the faraway package, you can type library(faraway) to
load the package and data(teengamb) to load the data. Useful R functions for 
this homework: `data(), summary(), hist(), plot(), boxplot()`. You can always 
type `help(subject)` to get detailed help on the subject, e.g. `help(plot)`. Or 
you can type `help.start()` to get interactive help with a search engine.

```{r hw1_source, message = FALSE}
source('./Stats500HW1.R')
```

#### Solution 1:
Below we can find a quick summary of the data. Since sex is a categorical 
variable, we are only given the number of observations of corresponding to the 
sex of the person. 

Observing the income and gamble variables, we notice that their means are 
greater than their medians. This is indicative that the distribution of the 
income and verbal variables are skewed right, i.e. greater portion of 
observations are bunched towards the left. The two histograms of these two 
variables below quickly demonstrate this explanation, with a large frequency of 
observations on the left.

Another connection between the income and gamble variables is that if we plot 
gamble against income, we can observe a positive linear trend arising. This 
linear trend would imply that the higher a person’s weekly income, the more 
they would be likely to spend on gambling per year. One may wonder if women 
would be more/less likely than men to spend on gambling. If we plot only male 
income observations and only female income observations against their 
corresponding gambling expenditures, we get some interesting results. Males 
seem to be driving this linear relationship the between income and gambling 
expenditures the most. Keep in mind that the y-axes of these two graphs are not 
the same, which demonstrates that females also spend a lot less on gambling 
than men do. 

#### Solution 2:

The description of the verbal variable from the `teengamb` data set defines the 
maximum score as 12. Looking at the histogram below for the verbal variable, we 
see that its values range from 1 – 10. The histogram also shows that the verbal 
score of 6 has the highest frequency. By inspection, the observations seem to 
follow a normal distribution with greater frequency about the middle and less 
frequency at the tails. 

The box plot of verbal easily shows that the verbal scores 1 and 2 are the 
outliers in the data. Running the command on R: 
“teengamb[teengamb$verbal <= 2, ],” our output below shows rows 31 and 35 from 
the data set, with their corresponding values. Interestingly these two rows are 
starkly different in their income and gamble values given their similarity in 
verbal score. 

























