---
title: "Homework 2"
author: "Israel Diego"
date: "February 2, 2016"
output:
  html_document:
    includes:
      in_header: ./Web_Materials/header.html
      after_body: ./Web_Materials/disqus.html
    toc: true
    toc_float: true
    theme: cerulean
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

### Part A 
The dataset `uswages` is drawn as a sample from the Current Population Survey in 1988.

**Hints:** Useful R functions: data(), lm(), summary(), residuals(), fitted(),
which.max(), mean(), median(), cor(), plot(). Note that the experience
variable has some negative values which most likely indicate missing data.
Those observations should be removed from the analysis.

**1.** Fit a regression model with weekly wages as the response and years of 
education and experience as predictors. Present the output.

*Regressing weekly wages on years of experience and education gives the 
following output:*

(Table)

*1% level. This indicates that a one year increase in education, on average will 
increase weekly wages by \$51.87. A one year increase in work experience will 
increase weekly wages by \$9.33. *

**2.** What percentage of variation in the response is explained by these
predictors? (Percentage variance explained is the same as coefficient
of determination).


**3.** Which observation has the largest (positive) residual? Give the case
number.

**4.** Compute the mean and median of the residuals. Explain what the
difference between the mean and the median indicates.

**5.** For two people with the same education and one year difference in
experience, what would be the difference in predicted weekly wages?

**6.** Compute the correlation of the residuals with the fitted values. Plot
residuals against fitted values. Explain the value of this correlation
using the geometric (projection) interpretation of least squares.

```{r hw1_source, message = FALSE}
source('./Stats500HW2.R')
```

### Part B
Using R, create a $10\times3$ matrix $X$:

\begin{align*}
X= & \left(\begin{array}{ccc}
1 & 2 & -2\\
1 & -1 & -2\\
1 & 3 & -2\\
1 & 3 & 3\\
1 & 2 & 3\\
1 & 1 & 3\\
1 & 0 & 0\\
1 & 0 & 0\\
1 & -1 & 0\\
1 & 0 & 1
\end{array}\right)
\end{align*}

Now create a $3\times1$ matrix $\beta$ whose entries are 1, -1,
and 2. Next create a $10\times1$ matrix $\epsilon$ whose entries
are IID standard normal (useful command: `rnorm`). Finally, set 
$Y=X\beta+\epsilon$.

**1.** Calculate $(X'X)^{-1}X'Y$ to estimate $\beta$. What do you get? (Don't use
the `lm` command. Do the computation directly. You can use the
`solve` command to compute a matrix inverse.)

**2.** What is the true variance of $\hat{\beta}$? (Remember that the variance of 
$\hat{\beta}$ is a $3\times3$ matrix.) (I say the "true" variance because, in this example, we know the true value of $\sigma^2$, and so don't need to estimate it 
using the residuals.)

**3.** Use the residuals to estimate $\sigma^2$. What do you get?

**4.** Now create a new $\epsilon$ and re-estimate $\beta$. Do this 1,000 times, 
and save all the answers in memory. Make a histogram of the 1,000 values of
$\hat{\beta_1}$. Do the same for $\hat{\beta_2}$ and $\hat{\beta_3}$. Also 
calculate the variance for each of these. Do your answers match with question 2?

**5.** Once again, re-create $\epsilon$ 1,000 times. Each time estimate $\beta$, 
and also estimate $\sigma^2$, too. Make a histogram of your 1,000 values of 
$\hat{\sigma^2}$. Based on the histogram, does it look like $\hat{\sigma^2}$ 
provides a reliable estimate of $\sigma^2$? Why do you think this is?

**6.** Repeat (4) and (5), but instead of using a normal distribution for 
$\epsilon$ use some other distribution that also has expectation 0 and variance 
1. Do your answers change much? Explain. You might want to experiment with a 
few different distributions.

























